\large{\bf Looking forwards and backwards through locally regulated populations }

### 2022-04-14 forwards-and-backwards.tex
comments on lookdown section

\plr{Motivation for and definition of the lookdown process; statement of main convergence theorem.}

@@

@@

A state of the discrete population model will be of the form $\xi=\sum\delta_{(x,u)}$,
where $x$ denotes the spatial location
of an individual, and $u$ is its level.
We shall abuse notation and treat $\xi$ both as a set and as a counting measure.
The spatial population described in \S \ref{sec: Model and main results} is recovered as
$$\eta=\frac{1}{\lambda}\sum_{(x,u)\in\xi}\delta_x,$$
i.e. $\eta$ is the re-scaled spatial projection of our spatial-level model $\xi$. 
Conditional on the spatial population $\eta$,
the levels of individuals in the population
will be independently and uniformly distributed
on $[0,\lambda]$.
Similar lookdown constructions can be found 
in \cite{etheridge/kurtz:2018},\cite{kurtz/rodrigues:2011}.
Test functions for this enriched process take the form
\begin{equation}
\label{test functions}
f(\xi)=\prod_{(x,u)\in \xi}g(x,u)=\exp\left(\int \log g(x,u)\xi(dx, du)\right),
\end{equation}
where $g(x,u)$ is differentiable in $u$ and 
smooth in $x$.
Moreover, $0\leq g(x,u) \leq 1$ for all $u\in [0,\lambda]$,
and $g(x,u)\equiv 1$ for $u\geq \lambda$.

@@

We should consider an initial population of $O(\lambda)$ many particles with levels
uniformly distributed between $[0, \lambda]$.
We write $(\xi(t))_{t \geq 0}$
for our population.
To ensure we have bounded intensity, 
we scale the empirical distribution of particles with a factor of $1/\lambda$,
so we write
$$\eta_t:=\frac{1}{\lambda}\xi_t(\cdot \times [0,\lambda])=
\frac{1}{\lambda}\sum\limits_{\substack{(X(t),U(t))\in \xi \\ U(t) < \lambda}} \delta_{X(t)}.$$

From here onwards,
we write $\xi$ to be the particle systems with levels,
while $\eta$ is the normalised empirical distribution.
An individual in the spatial population $\eta$ living at spatial location $x$
produces one adult offspring at rate $\gamma^{\mathfrak{m}}(x, \eta)$, where
$$\gamma^{\mathfrak{m}}(x,\eta)=\gamma(x,\eta) \int r (y, \eta)q(x,dy),$$
where $q(x,dy)$ is the dispersion kernel.

@@

See~\cite{etheridge/kurtz:2018}, Section~3.4 for a detailed explanation.
In order to incorporate death,
the level of the individual $(x,u)\in \eta$ 
moves upwards at an additional rate
$ \mu(x,\eta)$
(on reaching level $\lambda$
an individual is deemed to have died).
Overall, we assume that
the level associated with $(x,u)\in\eta$
evolves according to
$$\dot{u}= \gamma^{\mathfrak{m}}(x, \eta) 
\left\{\lambda^{-1}(\lambda -u)^{2}
-(\lambda -u)\right\}+\mu(x,\eta)u.$$
We shall write 
$$b(x,\eta):= \big\{\gamma^{\mathfrak{m}}(x,\eta)
-\mu(x,\eta)\big\}$$
to be the local branching rate.
Note that this quantity captures
the deviation from criticality
in the branching mechanism.
In this notation,
the differential equation 
governing the level of the individual $(x,u)\in\eta$ is
\begin{align}
\dot{u}&=\lambda  \gamma^{\mathfrak{m}}(x,\eta) 
\left\{\left(1-\frac{u}{\lambda}\right)^{2}
-1 +2\frac{u}{\lambda}\right\} -b(x,\eta)u \nonumber \\
&=\lambda^{-1}  \gamma^{\mathfrak{m}}(x,\eta) u^2 -b(x,\eta)u. \label{differential equation for level}
\end{align}
In this setting, the generator of our model
acting on the test functions $f(\xi)=\prod_{(x,u) \in \xi}g(x,u)$
satisfies
%We assume that  locations in ${\mathbb{ R}}^d$.

@@

We can now define our pre-limit explicitly as a solution to a Martingale Problem.
\begin{definition}[Martingale Problem Characterisation]
\label{Def: Rescaled Lookdown Models}

@@

>>>>

@@

\comment{Below, we write $\lambda$ instead of $N$: change to $N$?}

@@

A state of the discrete population model will be of the form $\xi=\sum\delta_{(x,u)}$,
where $x$ denotes the spatial location
of an individual, and $u$ is its level.
We shall abuse notation and treat $\xi$ both as a set and as a counting measure.
The spatial population described in \S \ref{sec: Model and main results} is recovered as
$$\eta=\frac{1}{\lambda}\sum_{(x,u)\in\xi}\delta_x,$$
i.e., $\eta$ is the re-scaled spatial projection of our spatial-level model $\xi$. 
Conditional on the spatial population $\eta$,
the levels of individuals in the population
will be independently and uniformly distributed
on $[0,\lambda]$.
Similar lookdown constructions can be found 
in \citet{etheridge/kurtz:2018} and \citet{kurtz/rodrigues:2011}.
Test functions for this enriched process take the form
\begin{equation}
\label{test functions}
f(\xi)=\prod_{(x,u)\in \xi}g(x,u)=\exp\left(\int \log g(x,u)\xi(dx, du)\right),
\end{equation}
where $g(x,u)$ is differentiable in $u$ and 
smooth in $x$.
We will also assume that $0\leq g(x,u) \leq 1$ for all $u\in [0,\lambda]$,
and $g(x,u)\equiv 1$ for $u\geq \lambda$.

@@

\comment{This repeats some stuff from above.}
We should consider an initial population of $O(\lambda)$ many particles with levels
uniformly distributed between $[0, \lambda]$.
We write $(\xi(t))_{t \geq 0}$
for our population.
To ensure we have bounded intensity, 
we scale the empirical distribution of particles with a factor of $1/\lambda$,
so we write
$$\eta_t:=\frac{1}{\lambda}\xi_t(\cdot \times [0,\lambda])=
\frac{1}{\lambda}\sum\limits_{\substack{(X(t),U(t))\in \xi \\ U(t) < \lambda}} \delta_{X(t)}.$$

From here onwards,
we write $\xi$ to be the particle systems with levels,
while $\eta$ is the normalised empirical distribution.
An individual in the spatial population $\eta$ living at spatial location $x$
produces one adult offspring at rate $\gamma^{\mathfrak{m}}(x, \eta)$, where
$$\gamma^{\mathfrak{m}}(x,\eta)=\gamma(x,\eta) \int r (y, \eta)q(x,dy),$$
where $q(x,dy)$ is the dispersal kernel.

\comment{Note: in what happens below, $\gamma^\mathfrak{m}$ is useful in that it makes the
equation for $\dot u$ nicer, but we could remove $q^\mathfrak{m}$ if we wanted
(although then \eqref{Generator of prelimit simplest model}
would then have both a $\gamma$ and a $\gamma^\mathfrak{m}$ in it).}

@@

See~\cite{etheridge/kurtz:2018}, Section~3.4 for a detailed explanation.
In order to incorporate death,
the level of the individual $(x,u)\in \eta$ 
moves upwards at an additional rate
$ \mu(x,\eta)$
(on reaching level $\lambda$
an individual is deemed to have died).
Overall, we assume that
the level associated with $(x,u)\in\eta$
evolves according to
$$\dot{u}= \gamma^{\mathfrak{m}}(x, \eta) 
\left\{\lambda^{-1}(\lambda -u)^{2}
-(\lambda -u)\right\}+\mu(x,\eta)u.$$
We shall write 
$$b(x,\eta):= \big\{\gamma^{\mathfrak{m}}(x,\eta)
-\mu(x,\eta)\big\}$$
to be the local branching rate.
Note that this quantity captures
the deviation from criticality
in the branching mechanism.
\comment{TODO: explain the connection to $F$.}
In this notation,
the differential equation 
governing the level of the individual $(x,u)\in\eta$ is
\comment{TODO: check that $\theta$ is right here, and consistent with the discussion above.}
\begin{align}
\dot{u}&=\lambda  \gamma^{\mathfrak{m}}(x,\eta) 
\left\{\left(1-\frac{u}{\lambda}\right)^{2}
-1 +2\frac{u}{\lambda}\right\} -b(x,\eta)u \nonumber \\
&=\lambda^{-1}  \gamma^{\mathfrak{m}}(x,\eta) u^2 -b(x,\eta)u. \label{differential equation for level}
\end{align}
In this setting, the generator of our model
acting on the test functions $f(\xi)=\prod_{(x,u) \in \xi}g(x,u)$
satisfies
%We assume that  locations in ${\mathbb{ R}}^d$.

@@

We can now define our pre-limit explicitly as a solution to a Martingale Problem.
\comment{TODO: this is a bit redundant with the above.
Instead, maybe (a) describe the process in words; (b) give this definition;
then (c) explain why it's the generator, so we only write down the generator once?}

% % % % % % % %
\begin{definition}[Martingale Problem Characterisation]
\label{Def: Rescaled Lookdown Models}

@@

\comment{Remark on the Remark: this is more than one remark.
I'm not sure it needs to be in a Remark block: how about just a few paragraphs?}
